---
title: "analysis_spring_health"
author: "Matthew Hanauer"
date: "4/18/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Step: Library packages and load data
```{r}
library(pacman)
p_load(dplyr, tidyverse, janitor, ggplot2)
data_spring_health <- read_csv("~/spring_health_assignment/spring_health_take_home_df.csv")
```
Step: Clean data sets
I am just getting all names to lower case snake case 
```{r}
data_spring_health_clean =  data_spring_health %>%
  clean_names() 


```


Step: Q1 How many individuals used our platform
Assuming the number of participants is the unique number of member_id_hashed.  It seems like each person takes both the SDS and PHQ-9
```{r}
number_of_participants = data_spring_health_clean %>%
  distinct(member_id_hashed) %>%
  count()

number_of_participants
```
Step: Q2 What is the average number of times that a member interacts with the platform.
Assuming member_id_hashed is the member id and is unique for each member.  So grouping them by member ID and then taking the mean.
```{r}
avg_number_of_interactions = data_spring_health_clean %>%
  group_by(member_id_hashed) %>%
  count() %>%
  ungroup() %>%
  summarise(mean_interactions = mean(n)) %>%
  mutate(mean_interactions = round(mean_interactions, 2))

avg_number_of_interactions
```
Step: Q3 What is the distribution of baseline PHQ9 total scores for members on the platform.
    Q3a Please include a visualization
    Q3b Please calculate summary statistics
```{r}
ggplot(data_spring_health_clean, aes(x=as.numeric(phq9_score))) + 
  geom_histogram(binwidth=1) + 
  labs(y= "Count of PHQ-9 total scores", x = "PHQ-9 total scores")

```

